\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2023-12-22 paper/main.tex
fixed typo

Consistent with the self-paced reading study, there is clear evidence for semantic interference, and evidence against syntactic interference.  How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory? We turn to this point next by discussing the quantiative predictions of the \textcite{Lewis2005} model.

@@

\textcite{mertzen} presented simulations that show approximately equally sized interference effects 
for both the syntactic and semantic manipulations. Their model assumed that three cues were used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}. The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations. \ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter \textit{F} and then generating data from the model while holding all other parameters in the model at fixed values. The \textit{F} parameter scales the activation to a retrieval time. The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020}. Figure \ref{fig:model_predictions_effects} shows the effects of interest computed based on the condition means shown in \ref{fig:model_predictions_rts}.

\begin{figure}[htbp]
    \centering
    \caption{Prior predictions from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}. 
    The implemented models are available with the code and data provided with this paper.}
    \label{fig:model_predictions}
    \centering
         \begin{subfigure}[t]{\textwidth}
         \centering
        \caption{Prior predicted by-condition retrieval time means. The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. Diamonds represent the means (numbers to the right) and blue dots show each simulated data point.}
         \label{fig:model_predictions_rts}
         \includegraphics[width=\textwidth]{paper/images/PriorPredictedMeans_interACTversions.png}
     \end{subfigure}

>>>>

Consistent with the self-paced reading study, there is clear evidence for semantic interference, and evidence against syntactic interference.  How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory? We turn to this point next by discussing the quantitative predictions of the \textcite{Lewis2005} model.

@@

\textcite{mertzen} presented simulations that show approximately equally sized interference effects 
for both the syntactic and semantic manipulations. Their model assumed that three cues were used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}. The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations. \ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter \textit{F} and then generating data from the model while holding all other parameters in the model at fixed values. The \textit{F} parameter scales the activation to a retrieval time. The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020,vasishth2020_abc}. Figure \ref{fig:model_predictions_effects} shows the effects of interest computed based on the condition means shown in \ref{fig:model_predictions_rts}.

\begin{figure}[htbp]
    \centering
    \caption{Prior predictions from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}. 
    The implemented models are available with the code and data provided with this paper.}
    \label{fig:model_predictions}
    \centering
         \begin{subfigure}[t]{\textwidth}
         \centering
        \caption{Prior predicted by-condition retrieval times. The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. Diamonds represent the means (numbers to the right) and blue dots show each simulated data point.}
         \label{fig:model_predictions_rts}
         \includegraphics[width=\textwidth]{paper/images/PriorPredictedMeans_interACTversions.png}
     \end{subfigure}
